### 2023-05-22 proofs/tex/alg-subtyping-lemmas.tex
Soundness and completeness of subtyping

    \item  \ruleref{\ottdruleAPVarLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$ .
    Then $[[uv a⁺ = ∅]]$, and $[[us]] = [[· : ·]]$ satisfies $[[us : Θ|∅]]$.
    Since $[[uv a⁺ = ∅]]$, application of any unification solution $[[us']]$ does not change $[[a⁺]]$, i.e.
    $[[ [us'] a⁺ = a⁺]]$. Therefore, $[[Γ ⊢ [us']a⁺ ≥ a⁺]]$ holds by \ruleref{\ottdruleDOneNVarLabel}.

    \item \label{case:pos-subt-soundness:arrow} 
     \ruleref{\ottdruleAShiftDLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ ↓uN ≥ ↓iM ⫤ us]]$.\\
    Then the next step of the algorithm is the unification of $[[nf(uN)]]$ and $[[nf(iM)]]$.
    By the soundness of the unification algorithm (\cref{lemma:unification-soundness}),
    it returns an equivalence-only solution $[[us]]$ such that $[[us : Ord uv uN]]$.
    By \cref{todo}, since $[[us]]$ is equivalence-only and $[[Γ ⊢ Θ]]$, $[[Θ ⊢ us' ⇒ us]]$ means 
    $[[ Γ ⊢ Sub us' ≈ Sub us : Ord uv uN ]]$ as substitutions.
    $[[ [us]nf(uN) = nf(iM) ]]$ implies $[[Γ ⊢ [us]nf(uN) ≈ nf(iM)]]$, and then 
        $[[Γ ⊢ [us']nf(uN) ≈ nf(iM)]]$. \ilyam{add lemmas}

@@

        Notice that $[[Γ, nbs ⊢ Θ, â⁻*[Γ, nbs] ]]$, $[[Γ,nbs ⊢ iQ']]$, and 
        $[[Γ,nbs; Θ, â⁻*[Γ, nbs] ⊢ [â⁻*/nas] uP' ]]$, so the induction hypothesis is applicable,
        that is $[[us' : Θ | uv [â⁻*/nas]uP']]$ and $[[ Γ, nbs ⊢ [us'2][â⁻*/nas]uP' ≥ iQ' ]]$ for any
        $[[us'2]]$ s.t. $[[Θ ⊢ us'2 ⇒ us']]$.

        Since the domain of $[[us']]$ is $[[uv [â⁺*/pas]uN']]$, the domain of 
        $[[us]] = [[us' \  {α̂⁺*}]]$ is $[[uv [â⁺*/pas]uN \ {â⁺*} = uv uN']] = [[uv ∀pas.uN']]$,
        this way, $[[us : Θ | uv ∀pas.uN']]$, as required.

        It is left to show that $[[Γ ⊢ [us2]∀pas.uN' ≤ ∀pbs.iM']]$ for any $[[us2]]$ s.t. $[[Θ ⊢ us2 ⇒ us]]$.
        Let us consider an arbitrary such $[[us2]]$. Let us construct $[[us'2]]$, 
        extending $[[us2]]$ to the domain $[[Ord uv [â⁺*/pas]uN']]$ with the values of $[[us']]$,
        i.e.  $[[us'2 : Θ | uv [â⁺*/pas]uN']]$, 
        and
        \[
            [[us'2(β̂±)]]  = 
            \begin{cases}
               [[us2(β̂±)]] & \text{if } [[β̂±]] \in [[uv uN']] \\
               [[us'(β̂±)]] & \text{if } [[β̂±]] \in [[â⁺*]]
            \end{cases}
        \] 
        , where the application of the unification solution to a variable returns the 
        corresponding \emph{unification entry}. 
        Notice that $[[Sub us'2|uv uN']] = [[us2]]$.
    It is easy to see that $[[Θ ⊢ us'2 ⇒ us']]$: 
    \begin{enumerate}
        \item if $[[β̂±]] \in [[uv uN']]$ then $[[us'2(β̂±)]] = [[us2(β̂±)]] \Rightarrow [[us(β̂±)]] = [[us'(β̂±)]]$;
        \item it $[[β̂±]] \in [[â⁺*]]$ then $[[us'2(β̂±)]] = [[us'(β̂±)]] \Rightarrow [[us'(β̂±)]]$,
    \end{enumerate}
    which means that the induction hypothesis can be applied to $[[us'2]]$, i.e.
    $[[ Γ, pbs ⊢ [us'2][â⁺*/pas]uN' ≤ iM' ]]$.\\
    Notice that
    $
    \begin{aligned}[t]
                 [[ [us'2][â⁺*/pas]uN' ]] &= [[ [Sub us'2|{â⁺*} ○ â⁺*/pas][Sub us'2|uv uN']uN' ]]
                                          && \text{by substitution properties \ilyam{todo}}\\
                                          &= [[ [Sub us'2|{â⁺*} ○ â⁺*/pas][us2]uN' ]]
                                          && \text{since $[[Sub us'2|uv uN']] = [[us2]]$}.
    \end{aligned}
    $\\
    Also notice that the domain of $[[Sub us'2|{â⁺*} ○ â⁺*/pas]]$ is $[[pas]]$,
    and the range is $[[Γ, pbs]]$, i.e. $[[Γ, pbs ⊢ Sub us'2|{â⁺*} ○ â⁺*/pas : pas]]$, 
    which means that we can apply \ruleref{\ottdruleDOneForallLabel} to 
    $[[ Γ, pbs ⊢ [Sub us'2|{â⁺*} ○ â⁺*/pas][us2]uN' ≤ iM' ]]$
    to get $[[ Γ ⊢ [us2]∀pas.uN' ≤ ∀pbs.iM' ]]$, as required.

\begin{lemma}[Soundness of Negative Subtyping] \label{lemma:neg-subt-soundness}
        If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iM]]$, and $[[Γ ; Θ ⊢ uN]]$ then\\ 
        $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$
        implies $[[us : Θ|uv uN]]$ and 
        for any $[[us']]$ such that $[[Θ ⊢ us' ⇒ us]]$,
        $[[ Γ ⊢ [us']uN ≤ iM ]]$
\end{lemma}
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$.
    Let us consider the last rule to infer this judgment. 
    \begin{caseof}
        \item $[[G;Θ ⊨ a⁻ ≤ a⁻ ⫤ ·]]$\\
        Then $[[uv a⁻ = ∅]]$, and $[[us]] = [[· : ·]]$ satisfies $[[us : Θ|∅]]$.
        Since $[[uv a⁻ = ∅]]$, application of any unification solution $[[us']]$ does not change $[[a⁻]]$, i.e.
        $[[ [us']a⁻ = a⁻ ]]$. Therefore, $[[Γ ⊢ [us']a⁻ ≤ a⁻]]$ holds by \ruleref{\ottdruleDOneNVarLabel}.

        \item $[[G;Θ ⊨ ↑uP ≤ ↑iQ ⫤ us]]$\\
        Then the next step of the algorithm is the unification of $[[nf(uP)]]$ and $[[nf(iQ)]]$.
        By the soundness of the unification algorithm (\cref{lemma:unification-soundness}),
        it returns an equivalence-only solution $[[us]]$ such that $[[us : Ord uv uP]]$.
        By \cref{todo}, since $[[us]]$ is equivalence-only and $[[Γ ⊢ Θ]]$, $[[Θ ⊢ us' ⇒ us]]$ means 
        $[[ Γ ⊢ Sub us' ≈ Sub us : Ord uv uP ]]$ as substitutions.

        $[[ [us]nf(uP) = nf(iQ) ]]$ implies $[[Γ ⊢ [us]nf(uP) ≈ nf(iQ)]]$, and then 
        $[[Γ ⊢ [us']nf(uP) ≈ nf(iQ)]]$. \ilyam{add lemmas}

        Let us rewrite the left-hand side and the right-hand side of $[[Γ ⊢ [us']nf(uP) ≈ nf(iQ)]]$ by 
        transitivity of equivalence (\cref{corollary:equivalence-transitivity}).
        By \cref{corollary:nf-sound-wrt-subt-equiv,corollary:subst-pres-equiv},
        $[[Γ ⊢ [us']nf(uP) ≈ [us']uP ]]$. By \cref{corollary:nf-sound-wrt-subt-equiv}, 
        $[[Γ ⊢ nf(iQ) ≈ iQ ]]$. 
        This way, we have $[[Γ ⊢ [us']uP ≈ iQ]]$.
        Then by \ruleref{\ottdruleDOneShiftULabel}
        and congruence of substitution, $[[Γ ⊢ [us']↑uP ≤ ↑iQ]]$.

        \item $[[G;Θ ⊨ uP → uN' ≤ iQ → iM' ⫤ us]]$\\
        The next step of the algorithm is two recursive calls:
        $[[G;Θ ⊨ uP ≥ iQ ⫤ us1]]$ and $[[G;Θ ⊨ uN' ≤ iM' ⫤ us2]]$.
        By the soundness of positive subtyping (\cref{lemma:pos-subt-soundness}) and induction hypothesis, respectively 
        \begin{enumerate}
            \item $[[us1 : Θ | uv uP]]$ and $[[ Γ ⊢ [us1']uP ≥ iQ ]]$ for any $[[us1']]$ s.t. $[[Θ ⊢ us1' ⇒ us1]]$
            \item $[[us2 : Θ | uv uN']]$ and $[[ Γ ⊢ [us2']uN' ≤ iM' ]]$ for any $[[us2']]$ s.t. $[[Θ ⊢ us2' ⇒ us2]]$
        \end{enumerate}

@@

        \item $[[G;Θ ⊨ ∀pas.uN' ≤ ∀pbs.iM' ⫤ us]]$ s.t. either $[[pas]]$ or $[[pbs]]$ is not empty\\
        Then the algorithm creates fresh unification variables $[[â⁺*[Γ,pbs] ]]$, 
        substitutes the old $[[pas]]$ with them in $[[uN']]$, and makes the recursive call:
        $[[G, pbs; Θ, â⁺*[G, pbs] ⊨ [â⁺*/pas] uN ≤ iM ⫤ us']]$, returning as the resulting solution 
        $[[us]] = [[us' \ {α̂⁺*}]]$.

        Notice that $[[Γ, pbs ⊢ Θ, â⁺*[G, pbs] ]]$, $[[Γ,pbs ⊢ iM']]$, and 
        $[[Γ,pbs; Θ, â⁺*[G, pbs] ⊢ [â⁺*/pas] uN' ]]$, so the induction hypothesis is applicable,
        that is $[[us' : Θ | uv [â⁺*/pas]uN']]$ and $[[ Γ, pbs ⊢ [us'2][â⁺*/pas]uN' ≤ iM' ]]$ for any
        $[[us'2]]$ s.t. $[[Θ ⊢ us'2 ⇒ us']]$.

        Since the domain of $[[us']]$ is $[[uv [â⁺*/pas]uN']]$, the domain of 
        $[[us]] = [[us' \  {α̂⁺*}]]$ is $[[uv [â⁺*/pas]uN \ {â⁺*} = uv uN']] = [[uv ∀pas.uN']]$,
        this way, $[[us : Θ | uv ∀pas.uN']]$, as required.

        It is left to show that $[[Γ ⊢ [us2]∀pas.uN' ≤ ∀pbs.iM']]$ for any $[[us2]]$ s.t. $[[Θ ⊢ us2 ⇒ us]]$.
        Let us consider an arbitrary such $[[us2]]$. Let us construct $[[us'2]]$, 
        extending $[[us2]]$ to the domain $[[Ord uv [â⁺*/pas]uN']]$ with the values of $[[us']]$,
        i.e.  $[[us'2 : Θ | uv [â⁺*/pas]uN']]$, 
        and
        \[
            [[us'2(β̂±)]]  = 
            \begin{cases}
               [[us2(β̂±)]] & \text{if } [[β̂±]] \in [[uv uN']] \\
               [[us'(β̂±)]] & \text{if } [[β̂±]] \in [[â⁺*]]
            \end{cases}
        \] 
        , where the application of the unification solution to a variable returns the 
        corresponding \emph{unification entry}. 
        Notice that $[[Sub us'2|uv uN']] = [[us2]]$.
    It is easy to see that $[[Θ ⊢ us'2 ⇒ us']]$: 
    \begin{enumerate}
        \item if $[[β̂±]] \in [[uv uN']]$ then $[[us'2(β̂±)]] = [[us2(β̂±)]] \Rightarrow [[us(β̂±)]] = [[us'(β̂±)]]$;
        \item it $[[β̂±]] \in [[â⁺*]]$ then $[[us'2(β̂±)]] = [[us'(β̂±)]] \Rightarrow [[us'(β̂±)]]$,
    \end{enumerate}
    which means that the induction hypothesis can be applied to $[[us'2]]$, i.e.
    $[[ Γ, pbs ⊢ [us'2][â⁺*/pas]uN' ≤ iM' ]]$.\\
    Notice that
    $
    \begin{aligned}[t]
                 [[ [us'2][â⁺*/pas]uN' ]] &= [[ [Sub us'2|{â⁺*} ○ â⁺*/pas][Sub us'2|uv uN']uN' ]]
                                          && \text{by substitution properties \ilyam{todo}}\\
                                          &= [[ [Sub us'2|{â⁺*} ○ â⁺*/pas][us2]uN' ]]
                                          && \text{since $[[Sub us'2|uv uN']] = [[us2]]$}.
    \end{aligned}
    $\\
    Also notice that the domain of $[[Sub us'2|{â⁺*} ○ â⁺*/pas]]$ is $[[pas]]$,
    and the range is $[[Γ, pbs]]$, i.e. $[[Γ, pbs ⊢ Sub us'2|{â⁺*} ○ â⁺*/pas : pas]]$, 
    which means that we can apply \ruleref{\ottdruleDOneForallLabel} to 
    $[[ Γ, pbs ⊢ [Sub us'2|{â⁺*} ○ â⁺*/pas][us2]uN' ≤ iM' ]]$
    to get $[[ Γ ⊢ [us2]∀pas.uN' ≤ ∀pbs.iM' ]]$, as required.

@@

    \begin{caseof}
        \item \label{case:subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ (for potentially empty $[[pbs]]$)\\
        Then by assumption, there exists $[[us : Θ | uv uP]]$ such that $[[ Γ ⊢ ∃nbs.[us]α̂⁺ ≥ iQ ]]$.
        $[[us : Θ | uv α̂⁺]]$ means that  $[[us]]$ is either $[[ (α̂⁺ :≈ iP') ]]$ or $[[ (α̂⁺ :≥ iP') ]]$,
        where $[[Δ ⊢ iP']]$ and $[[α̂⁺[Δ] ∊ Θ]]$.
        $[[ Γ ⊢ ∃nbs.[us]α̂⁺ ≥ iQ ]]$ means that $[[Γ ⊢ iP' ≥ iQ]]$
        because multiple inversions of \ruleref{\ottdruleDOneExistsLabel} 
        gives us $[[Γ ⊢ [us]α̂⁺ ≥ iQ]]$ since $[[ {nbs} ∩ fv [us]α̂⁺ = ∅]]$.

@@

        \item \label{case:subt-complete-nvar}
        $[[ Γ ⊢ [us]uN ≤ iM ]]$ is derived by \ruleref{\ottdruleDOneNVarLabel}, 
        i.e. $[[ [us]uN ]] = [[iM]] = [[ α⁻ ]] = [[iN]]$ 
        (the latter equality holds because $[[uN]] \neq [[α̂⁻]]$).

        Notice that $[[us : Θ | uv α⁻]]$ = $[[us : Θ | ∅]]$ = $[[·]]$.

        The algorithm applies \ruleref{\ottdruleANVarLabel} and 
        infers $[[us']] = [[·]]$, i.e. $[[Γ;Θ ⊨ a⁻ ≤ a⁻ ⫤ ·]]$.

@@

       \item The positive case when $[[Γ ⊢ [us]uP ≥ iQ]]$ is derived by 
       \ruleref{\ottdruleDOneShiftDLabel} is symmetric to the corresponding negative
        \cref{case:subt-complete-upshift}. Notice that the algorithm does not 
        make a recursive call in this case but rather invoke the unification.
        This way, the proof does not apply the induction hypothesis, and hence, 
        the absence of negative unification variables in a negative type
        ($[[α̂⁻]] \notin [[uN]]$ for $[[↓uN]] =[[uP]]$) is not required.

      \item The positive case when $[[Γ ⊢ [us]uP ≥ iQ]]$ is derived by 
      \ruleref{\ottdruleDOneExistsLabel} is symmetric to the corresponding negative
      \cref{case:subt-complete-forall}. Notice that we only consider the case
      when the substitution $[[ [us]uP ]]$ results in the existential type 
      $[[∃nas.iP']]$ (for $[[iP']] \neq [[∃]]\dots$) by congruence, 
      i.e. $[[uP = ∃nas.uP'']]$ (for $[[uP'']] \neq [[∃]]\dots$) and $[[ [us]uP'' = iP' ]]$.
      This is because the case when $[[uP = ∃nbs.α̂⁺]]$ has been covered
      (\cref{case:subt-complete-base}), and thus, the substitution $[[us]]$ must
      preserve all the outer quantifiers of $[[uP]]$ and does not generate any new ones.

>>>>

    \item \label{case:pos-subt-soundness:var} \ruleref{\ottdruleAPVarLabel}  
    then $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$ .
    Then $[[uv a⁺ = ∅]]$, and $[[us]] = [[· : ·]]$ satisfies $[[us : Θ|∅]]$.
    Since $[[uv a⁺ = ∅]]$, application of any unification solution $[[us']]$ does not change $[[a⁺]]$, i.e.
    $[[ [us'] a⁺ = a⁺]]$. Therefore, $[[Γ ⊢ [us']a⁺ ≥ a⁺]]$ holds by \ruleref{\ottdruleDOneNVarLabel}.

    \item \label{case:pos-subt-soundness:shift} 
     \ruleref{\ottdruleAShiftDLabel} then
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ us]]$ has shape $[[Γ;Θ ⊨ ↓uN ≥ ↓iM ⫤ us]]$.\\
    Then the next step of the algorithm is the unification of $[[nf(uN)]]$ and $[[nf(iM)]]$.
    By the soundness of the unification algorithm (\cref{lemma:unification-soundness}),
    it returns an equivalence-only solution $[[us]]$ such that $[[us : Ord uv uN]]$.
    By \cref{todo}, since $[[us]]$ is equivalence-only and $[[Γ ⊢ Θ]]$, $[[Θ ⊢ us' ⇒ us]]$ means 
    $[[ Γ ⊢ Sub us' ≈ Sub us : Ord uv uN ]]$ as substitutions.
    $[[ [us]nf(uN) = nf(iM) ]]$ implies $[[Γ ⊢ [us]nf(uN) ≈ nf(iM)]]$, and then 
        $[[Γ ⊢ [us']nf(uN) ≈ nf(iM)]]$. \ilyam{add lemmas}

@@

        Notice that $[[Γ, nbs ⊢ Θ, â⁻*[Γ, nbs] ]]$, $[[Γ,nbs ⊢ iQ']]$, and 
        $[[Γ,nbs; Θ, â⁻*[Γ, nbs] ⊢ [â⁻*/nas] uP' ]]$, so the induction hypothesis is applicable,
        that is $[[us' : Θ, â⁻*[Γ, nbs] | uv [â⁻*/nas]uP']]$ and $[[ Γ, nbs ⊢ [us'2][â⁻*/nas]uP' ≥ iQ' ]]$ for any
        $[[us'2]]$ s.t. $[[Θ, â⁻*[Γ, nbs] ⊢ us'2 ⇒ us']]$.

        Since the domain of $[[us']]$ is $[[uv [â⁻*/nas]uP']]$, the domain of 
        $[[us]] = [[us' \  {α̂⁻*}]]$ is $[[uv [â⁻*/nas]uP' \ {â⁻*}]] = [[uv ∃nas.uP']]$,
        this way, $[[us : Θ | uv ∃nas.uP']]$, as required.

        It is left to show that $[[Γ ⊢ [us2]∃nas.uP' ≥ ∃nbs.iQ']]$ for any $[[us2]]$ s.t. $[[Θ ⊢ us2 ⇒ us]]$.
        Let us consider an arbitrary such $[[us2]]$. Let us construct $[[us'2]]$, 
        extending $[[us2]]$ to the domain $[[uv [â⁻*/nas]uP']]$ with the values of $[[us']]$,
        i.e.  $[[us'2 : Θ | uv [â⁻*/nas]uP']]$, 
        and
        \[
            [[us'2(β̂±)]]  = 
            \begin{cases}
               [[us2(β̂±)]] & \text{if } [[β̂±]] \in [[uv uP']] \\
               [[us'(β̂±)]] & \text{if } [[β̂±]] \in [[â⁻*]]
            \end{cases}
        \] 
        , where the application of the unification solution to a variable returns the 
        corresponding \emph{unification entry}. 
        Notice that $[[Sub us'2|uv uP']] = [[us2]]$.
    It is easy to see that $[[Θ ⊢ us'2 ⇒ us']]$: 
    \begin{enumerate}
        \item if $[[β̂±]] \in [[uv uP']]$ then $[[us'2(β̂±)]] = [[us2(β̂±)]] \Rightarrow [[us(β̂±)]] = [[us'(β̂±)]]$;
        \item it $[[β̂±]] \in [[â⁻*]]$ then $[[us'2(β̂±)]] = [[us'(β̂±)]] \Rightarrow [[us'(β̂±)]]$,
    \end{enumerate}
    which means that the induction hypothesis can be applied to $[[us'2]]$, i.e.
    $[[ Γ, nbs ⊢ [us'2][â⁻*/nas]uP' ≥ iQ' ]]$.\\
    Notice that
    $
    \begin{aligned}[t]
                 [[ [us'2][â⁻*/nas]uP' ]] &= [[ [Sub us'2|{â⁻*} ○ â⁻*/nas][Sub us'2|uv uP']uP' ]]
                                          && \text{by substitution properties \ilyam{todo}}\\
                                          &= [[ [Sub us'2|{â⁻*} ○ â⁻*/nas][us2]uP' ]]
                                          && \text{since $[[Sub us'2|uv uP']] = [[us2]]$}.
    \end{aligned}
    $\\
    Also notice that the domain of $[[Sub us'2|{â⁻*} ○ â⁻*/nas]]$ is $[[nas]]$,
    and the range is $[[Γ, nbs]]$, i.e. $[[Γ, nbs ⊢ Sub us'2|{â⁻*} ○ â⁻*/nas : nas]]$, 
    which means that we can apply \ruleref{\ottdruleDOneForallLabel} to 
    $[[ Γ, nbs ⊢ [Sub us'2|{â⁻*} ○ â⁻*/nas][us2]uN' ≤ iM' ]]$
    to obtain $[[ Γ ⊢ [us2]∃nas.uP' ≥ ∃nbs.iQ' ]]$, as required.
    \end{caseof}
\end{proof}

\begin{lemma}[Soundness of Negative Subtyping] \label{lemma:neg-subt-soundness}
        If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iM]]$, and $[[Γ ; Θ ⊢ uN]]$ then\\ 
        $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$
        implies $[[us : Θ|uv uN]]$ and 
        for any $[[us']]$ such that $[[Θ ⊢ us' ⇒ us]]$,
        $[[ Γ ⊢ [us']uN ≤ iM ]]$
\end{lemma}
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$.
    Let us consider the last rule to infer this judgment. 
    \begin{caseof}
        \item \ruleref{\ottdruleAArrowLabel}, and then $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$ has shape
        $[[G;Θ ⊨ uP → uN' ≤ iQ → iM' ⫤ us]]$\\
        On the next step, the the algorithm makes two recursive calls:
        $[[G;Θ ⊨ uP ≥ iQ ⫤ us1]]$ and $[[G;Θ ⊨ uN' ≤ iM' ⫤ us2]]$.
        By the soundness of positive subtyping (\cref{lemma:pos-subt-soundness}) and induction hypothesis, respectively 
        \begin{enumerate}
            \item $[[us1 : Θ | uv uP]]$ and $[[ Γ ⊢ [us1']uP ≥ iQ ]]$ for any $[[us1']]$ s.t. $[[Θ ⊢ us1' ⇒ us1]]$
            \item $[[us2 : Θ | uv uN']]$ and $[[ Γ ⊢ [us2']uN' ≤ iM' ]]$ for any $[[us2']]$ s.t. $[[Θ ⊢ us2' ⇒ us2]]$
        \end{enumerate}

@@

        \item \ruleref{\ottdruleANVarLabel}, and then $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$ has shape $[[G;Θ ⊨ a⁻ ≤ a⁻ ⫤ ·]]$\\
        This case is symmetric to \cref{case:pos-subt-soundness:var} of \cref{lemma:pos-subt-soundness}.

        \item \ruleref{\ottdruleAShiftULabel}, and then $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$ has shape $[[G;Θ ⊨ ↑uP ≤ ↑iQ ⫤ us]]$\\
        This case is symmetric to \cref{case:pos-subt-soundness:shift} of \cref{lemma:pos-subt-soundness}.

        \item \ruleref{\ottdruleAForallLabel}, and then $[[Γ ; Θ ⊨ uN ≤ iM ⫤ us]]$ has shape $[[G;Θ ⊨ ∀pas.uN' ≤ ∀pbs.iM' ⫤ us]]$ s.t. either $[[pas]]$ or $[[pbs]]$ is not empty\\
        This case is symmetric to \cref{case:pos-subt-soundness:exists} of \cref{lemma:pos-subt-soundness}.

@@

    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ (for potentially empty $[[pbs]]$)\\
        Then by assumption, there exists $[[us : Θ | uv uP]]$ such that $[[ Γ ⊢ ∃nbs.[us]α̂⁺ ≥ iQ ]]$.
        $[[us : Θ | uv α̂⁺]]$ means that  $[[us]]$ is either $[[ (α̂⁺ :≈ iP') ]]$ or $[[ (α̂⁺ :≥ iP') ]]$,
        where $[[Δ ⊢ iP']]$ and $[[α̂⁺[Δ] ∊ Θ]]$.
        $[[ Γ ⊢ ∃nbs.[us]α̂⁺ ≥ iQ ]]$ means that $[[Γ ⊢ iP' ≥ iQ]]$
        because multiple inversions of \ruleref{\ottdruleDOneExistsLabel} 
        gives us $[[Γ ⊢ [us]α̂⁺ ≥ iQ]]$ since $[[ {nbs} ∩ fv [us]α̂⁺ = ∅]]$.

@@

        \item \label{case:pos-subt-complete-pvar}
        $[[ Γ ⊢ [us]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOnePVarLabel}, 
        i.e. $[[iP]] = [[ [us]uP ]] = [[ α⁺ ]] = [[iQ]]$.
        Here the first equality holds because $[[uP]]$ is not a unification variable:
        this case has been covered by \cref{case:pos-subt-complete-base}.
        The second equality hold because \ruleref{\ottdruleDOnePVarLabel} was applied.

        Notice that $[[us : Θ | uv α⁺]]$ = $[[us : Θ | ∅]]$ = $[[·]]$.

        The algorithm applies \ruleref{\ottdruleAPVarLabel} and 
        infers $[[us']] = [[·]]$, i.e. $[[Γ;Θ ⊨ a⁺ ≥ a⁺ ⫤ ·]]$.

@@

        \item \label{case:pos-subt-complete-upshift} $[[ Γ ⊢ [us]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOneShiftDLabel},

        Then $[[ uP ]] = [[ ↓uN ]]$, since the substitution $[[ [us]uP ]]$ must preserve the 
        top-level constructor of $[[uP]]\neq [[α̂⁺]]$ (the case $[[uP]] = [[α̂⁺]]$ has been covered
        by \cref{case:pos-subt-complete-base}), and $[[uQ]] = [[ ↓iM ]]$,
        and by inversion, $[[ Γ ⊢ [us]uN ≈ iM ]]$.

        Since both types start with $[[↓]]$, 
        the algorithm tries to apply \ruleref{\ottdruleAShiftDLabel}: 
        $[[G;Θ ⊨ ↓uN ≥ ↓iM ⫤ us']]$. The premise of this rule is the
        unification of $[[nf(uN)]]$ and $[[nf(iM)]]$:
        $[[Γ;Θ ⊨ nf(uN) ≈u nf(iM) ⫤ us']]$. Let us show that the unification successfully 
        terminates and returns $[[us']] = [[nf(us)]]$.

        To demonstrate that the unification terminates, we apply the completeness 
        of the unification algorithm (\cref{lemma:unification-completeness}). 
        In order to do that, we need to provide a unifier of 
        $[[nf(uN)]]$ and $[[nf(iM)]]$. Thankfully, $[[nf(us)]]$ does it.

        \begin{itemize}
            \item $[[nf(uN)]]$ and $[[nf(iM)]]$ are normalized 
            \item $[[Γ ; Θ ⊢ nf(uN)]]$ because $[[Γ ; Θ ⊢ uN]]$ (\cref{todo})
            \item $[[Γ ⊢ nf(iM)]]$ because $[[Γ ⊢ iM]]$ (\cref{corollary:wf-nf})
            \item $[[nf(us) : Θ | uv nf(uN)]]$ because $[[us: Θ | uv uN]]$ (\cref{todo})
            \item $ \begin{aligned}[t]
                    [[ Γ ⊢ [us]uN ≈ iM ]] &\Rightarrow [[ [us]uN ≈ iM ]]
                                          && \text {by \cref{lemma:equiv-completeness}}\\
                                          &\Rightarrow [[ nf([us]uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:normalization-completeness}}\\
                                          &\Rightarrow [[ [nf(us)]nf(uN) = nf(iM) ]]
                                          && \text {by \cref{lemma:norm-subst-distr}}\\
                    \end{aligned}
                  $
        \end{itemize}
        Then by the completeness of the unification,
        $[[Γ ; Θ ⊨ nf(uN) ≈u nf(iM) ⫤ nf(us)]]$.
        This way, the subtyping algorithm terminates and the resulting solution is
        $[[nf(us)]]$.

      \item $[[ Γ ⊢ [us]uP ≥ iQ ]]$ is derived by \ruleref{\ottdruleDOneExistsLabel}.\\
      We should only consider the case
      when the substitution $[[ [us]uP ]]$ results in the existential type 
      $[[∃nas.iP'']]$ (for $[[iP'']] \neq [[∃]]\dots$) by congruence, 
      i.e. $[[uP = ∃nas.uP']]$ (for $[[uP']] \neq [[∃]]\dots$) and $[[ [us]uP' = iP'' ]]$.
      This is because the case when $[[uP = ∃nbs.α̂⁺]]$ has been covered
      (\cref{case:pos-subt-complete-base}), and thus, the substitution $[[us]]$ must
      preserve all the outer quantifiers of $[[uP]]$ and does not generate any new ones.

      This way, $[[uP]] = [[∃nas.uP']]$, $[[ [us]uP ]] = [[ ∃nas.[us]uP' ]]$ 
      (assuming $[[nas]]$ does not intersect with the range of $[[us]]$)
      and $[[iQ]] = [[ ∃nbs.iQ' ]]$, where either $[[nas]]$ or $[[nbs]]$ is not empty.

      By inversion, $[[ Γ ⊢ [σ][us]uP' ≥ iQ' ]]$ for some $[[Γ, nbs ⊢ σ : nas]]$.
      Since $[[σ]]$ and $[[us]]$ have disjoint domains,
      and the range of one does not intersect with the domain of the other,
      they commute, i.e. $[[ Γ, nbs ⊢ [us][σ]uP' ≥ iQ' ]]$
      (notice that the tree inferring this judgement is 
      a proper subtree of the tree inferring 
      $[[ Γ ⊢ [us]uP ≥ iQ ]]$).

      At the next step, 
      the algorithm creates fresh (disjoint with $[[uv uP']]$) 
      unification variables $[[â⁻*]]$, replaces $[[nas]]$ with them in $[[ uP' ]]$,
      and makes the recursive call:
      $[[G, nbs; Θ, â⁻*[G, nbs] ⊨ uP0 ≥ iQ ⫤ us1]]$,
      (where $[[uP0]] = [[ [â⁻*/nas]uP' ]]$),
      returning $[[us1 \ {â⁻*}]]$ as the result.

       Notice that $[[ [us][σ][nas/â⁻*]uP0 = [us][σ]uP' ]]$,
       and thus, $[[ Γ, nbs ⊢ [us][σ][nas/â⁻*]uP0 ≥ iQ' ]]$.
       Let us combine $[[us]]$ and $[[σ  ○ nas/â⁻*]]$ into one 
       $[[us0]]$---a unification solution for $[[uP0]]$:
        \[
            [[us0(β̂±)]]  = 
            \begin{cases}
               [[ β̂± :≈ [σ]αi⁻ ]] & \text{if } [[β̂±]] = [[αî⁻]] \in [[â⁻*]] \\
               [[us(β̂±)]] & \text{if } [[β̂±]] \in [[uv uP']]
            \end{cases}
       \]

       Let us show that the induction hypothesis is applicable to 
       $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ' ]]$, with the meta-context $[[ Θ, â⁻*[Γ, nbs] ]]$
       \begin{itemize}
        \item $[[ us0 : (Θ, â⁻*[Γ, nbs])|uv uP0 ]]$. Notice that for every $[[αî⁻]] \in [[â⁻*]]$, 
            the type corresponding to the entry $[[us0(αî⁻)]]$ 
            is well-formed in $[[Γ, nbs]]$ since $[[Γ, nbs ⊢ σ : nas ]]$;
            and for every $[[β̂±]] \in [[uv uP']]$, 
            the type corresponding to the entry $[[us0(β̂±)]]$ is well-formed in
            context $[[Θ(β̂±)]]$ since $[[us : Θ | uv uP']]$.
        \item $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ' ]]$. This is because
            $[[ [us0]uP0 ]] = [[ [us][σ][nas/â⁻*]uP0 ]] = [[ [us][σ]uP' ]]$.
        \item $[[ Γ, nbs ⊢ Θ, â⁻*[Γ, nbs] ]]$ since
            $[[Γ, nbs ⊢ Θ]]$ and $[[ {Γ, nbs} ⊆ {Γ, nbs} ]]$.
        \item $[[Γ, nbs ; Θ, â⁻*[Γ, nbs] ⊢ uP0 ]]$
       \end{itemize}

       This way, we apply the induction hypothesis to $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ' ]]$ and 
       infer that there exists\\
        $[[us0' :  (Θ, â⁻*[Γ, nbs]) | uv uP0]]$ such that
       $[[G, nbs; Θ, â⁻*[G, nbs] ⊨ uP0 ≥ iQ ⫤ us0']]$,
       and $[[Θ, â⁻*[G, nbs] ⊢ us0' ⇒ us0]]$.
